---
title: "Predicting the Quantified Self"
author: "Bastiaan Quast"
date: "Monday, July 21, 2014"
output:
  html_document:
    keep_md: yes
---

## Abstract
We use the random forest method to estimate features for the Human Activity Recognition data set from Groupware. We that this method produces relevant results.


## Data
The data is taken from the [Human Activity Recognition](http://groupware.les.inf.puc-rio.br/har) programme at [Groupware](http://groupware.les.inf.puc-rio.br/).

We start the data loading procedure by specifying the data sources and destinations.
```{r source.dest}
training.file <- 'pml-training.csv'
test.file     <- 'pml-test.csv'
training.url  <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.url      <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
```
We then execute the downloads.
```{r download, eval=FALSE}
download.file(training.url, training.file, method='wget')
download.file(test.url,     test.file,     method='wget')
```
In order to ensure all transformations are applied to both the trainig and test data sets equally, we employ the OOP method and create functions for every step. First we create a function to read data and set NAs and apply this to both datasets.
```{r read}
read.pml       <- function(x) { read.csv(x, na.strings = c("", "NA", "#DIV/0!") ) }
training       <- read.pml(training.file)
test           <- read.pml(test.file)
training       <- training[,-c(1,5,6)]
test           <- test[,-c(1,5,6)]
```

```{r split, message=FALSE}
library(caret)
trainingIndex  <- createDataPartition(training$classe, p=.50, list=FALSE)
training.train <- training[ trainingIndex,]
training.test  <- training[-trainingIndex,]
```

Next we create a function to remove entire NA columns, and apply it to both data frames. Lastly we create a function that removes any variables with missing NAs and apply this to the training set.
```{r nacols}
rm.na.cols     <- function(x) { x[ , colSums( is.na(x) ) < nrow(x) ] }
training.train <- rm.na.cols(training.train)
training.test  <- rm.na.cols(training.test)
complete       <- function(x) {x[,sapply(x, function(y) !any(is.na(y)))] }
incompl        <- function(x) {names( x[,sapply(x, function(y) any(is.na(y)))] ) }
trtr.na.var    <- incompl(training.train)
trts.na.var    <- incompl(training.test)
training.train <- complete(training.train)
training.test  <- complete(training.test)
```

## Method
We use the **Random Forests** method, which applies **bagging** to **tree learners**.

The **Random Forests** method is described by Leo Breiman in [Random Forests](http://dx.doi.org/10.1023%2FA%3A1010933404324).

The **B**ootstap **Agg**regat**ing** (bagging) method is described in [Technical Report No. 421: Bagging Predictors](http://dx.doi.org/10.1007%2FBF00058655) by Leo Breiman.
```{r train}
library(randomForest)
random.forest <- train(training.train[,-57],
                       training.train$classe,
                       tuneGrid=data.frame(mtry=3),
                       trControl=trainControl(method="none")
                       )
```


## Results
Some statistics on the results
```{r summary}
summary(random.forest)
```
We now compare the results from the predition with the actual data.
```{r confusion_matrix}
confusionMatrix(predict(random.forest,
                        newdata=training.test[,-57]),
                training.test$classe
                )
```
The Kappa statistic of 0.994 reflects the out-of-sample error.

```{r plot}
plot( varImp(random.forest) )
```